---
title: "input csvs creator"
author: "Chris Day"
date: "10/23/2021"
output: html_document
---

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(magrittr)
library(sf)
library(leaflet)
library(R.utils)
library(kableExtra)
```

## Introduction
In this RMarkdown, we create the taz-centroids.csv, taz-parking.csv, and tpcm-loc.csv files for the slc-light scenarios. The taz-centroids.csv file will have all the TAZ's ids, the x and y coordinate to its centorid, and the area. THe taz-parking.csv file will give basic parking values to each TAZ and give unlimiting parking to each one. Finally, the tpcm-loc.csv file is created to store important location variables used in the tour mode choice model. Specifically, TAZ location variables like whether it is a CBD area, as well as zonal density information will be included. 

## taz-centroids.csv

Here we create the taz-centroids.csv file. This file is important for BEAM to be able to guesstimate what taz a specific agent is in. 

```{r, message = FALSE, warning = FALSE}
# read in all the tazs into correct coordinate system
taz <- st_read("https://raw.githubusercontent.com/byu-transpolab/populationsim_wfrc/master/inputs/taz.geojson") %>% 
  st_transform("+init=epsg:26912") %>% ## correct crs???? CHECK THIS
  arrange(TAZID) %>% mutate(TAZ = as.character(TAZID)) %>%
  
  # remove external stations
  filter(CO_FIPS > 0) %>%
  select(TAZ, CO_NAME, PRKCSTPERM, PRKCSTTEMP, CBD, X, Y)

  # add centroid  and area geometry
  taz$geometry2 = st_centroid(st_geometry(taz))
  taz$area <- st_area(taz)
  
tazc <- taz
st_geometry(tazc) <- "geometry2"
```

Below is a map displaying all the tazs along with their centroid.

```{r, message = FALSE, warning = FALSE}
# display map of centroids and tazs
leaflet() %>%
  addTiles() %>%
  addPolygons(data = taz %>% st_transform(4326), color = "red") %>%
  addCircles(data = tazc %>% st_transform(4326))
```

Finally, we create the csv file that is need in BEAM. The table is also shown below. 
```{r, echo = FALSE}
# create taz-centroids.csv file
tazcsv <- taz %>% as_tibble() %>%
  mutate(taz = TAZ, 'coord-x' = X, 'coord-y' = Y) %>%
  select(taz, 'coord-x', 'coord-y', area)
#write_csv(tazcsv, "utah-taz-centers.csv")
kbl(head(tazcsv), booktabs = T) %>%
  kable_styling(latex_options = "striped")

```

## taz-parking.csv

In addition, the taz-parking.csv file needs to be created so that BEAM doesn't assign a default value of $50 to each agent parking! Below, that table is created.

```{r, message = FALSE, warning = FALSE}
# create default taz-parking csv using temperary parking costs
tazparking <- taz %>% as_tibble() %>%
  select(TAZ,PRKCSTPERM,PRKCSTTEMP) %>%
  mutate(
    taz = TAZ,
    parkingType = list(c("Public","Residential","Workplace")),
    pricingModel = "Block",
    chargingType = ifelse(taz %in% c("2870","2871","2872","2873"),"Level1(2.3|AC)", "NoCharger"),
    numStalls = 1000000,
    feeInCents = 100 * PRKCSTTEMP,
    reservedFor = "Any") %>%
  unnest(cols = c(parkingType)) %>%
  select(-TAZ, -PRKCSTPERM, -PRKCSTTEMP)
```

A section of that table is displayed below. 

```{r, echo = FALSE}
# write taz-parking csv
#write_csv(tazparking,"utah-taz-parking.csv")
kbl(head(tazparking), booktabs = T) %>%
  kable_styling(latex_options = "striped")
```

## tpcm-loc.csv

Finally, I create the location data file per taz. This is useful for calculating the utility for the mode choice model. 

```{r, message = FALSE, warning = FALSE}
# read in se data that has location variables per taz
sedata <- read_csv("https://raw.githubusercontent.com/byu-transpolab/populationsim_wfrc/master/SE_Data.csv") %>%
  mutate(TAZ = as.character(ZONE)) %>%
  select(TAZ, TOTHH, RESACRE, CIACRE, TOTEMP)
```

```{r, message = FALSE, warning = FALSE}
# combine with taz geojson file and select values we need
location <- taz %>% as_tibble() %>% 
  left_join(sedata, by = "TAZ") %>%
  mutate(
    tazid = TAZ, 
    tothh = as.numeric(TOTHH),
    resacre = as.numeric(RESACRE),
    ciacre = as.numeric(CIACRE),
    totemp = as.numeric(TOTEMP),
    cbd = as.numeric(CBD)
  ) %>%
  select(tazid, tothh, resacre, ciacre, totemp, cbd)

# create tpcm-loc csv file
#write_csv(location, "utah-tpcm-loc.csv")
```

Below, a section of this table is displayed. 
```{r, echo = FALSE}
kbl(head(location), booktabs = T) %>%
  kable_styling(latex_options = "striped")
```
 
